---
title: "| RESEARCH PROTOCOL\n| \n| Is fluoroquinolone use associated with the development of aortic aneurysms and aortic dissections \n"
fontsize: 12pt
geometry: margin=1in
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs") })
output:
  bookdown::html_document2:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float: yes
    number_sections: yes
    number_tables: yes
    css: style.css
  word_document:
    reference_docx: ohdsi-protocol-style.docx
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: xelatex
    md_extensions: +raw_attribute
    number_sections: yes
    includes:
      before_body: title.tex
header-includes:
- \usepackage[numbers,sort&compress]{natbib}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{caption}
- \usepackage{rotating}
- \usepackage{multirow}
- \usepackage{mwe,tikz}
- \usepackage[percent]{overpic}
- \usepackage{enumitem}
- \usepackage{hyperref}
- \newcolumntype{P}[1]{>{\raggedright\arraybackslash}p{#1}}
- \newcommand{\footerDate}{`r params$date`}
- \input{header.tex}
longtable: yes
mainfont: Arial
bibliography: Protocol.bib
params:
  date: '02-May-2023'
  version: 1.1.0
subtitle: 'Version: `r params$version`'
link-citations: yes
csl: bmj.csl
---

```{r setup, include=FALSE}
# library(LegendT2dm)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
library(dplyr)
options(readr.show_col_types = FALSE)
options(knitr.kable.NA = "")
options(knitr.table.format = function() {
  if (knitr::is_latex_output()) {
    "latex"
  } else if (knitr::is_html_output()) {
    "html"
  } else {
    "pipe"
  }
})
latex_table_font_size <- 8
```

# List of Abbreviations

```{r abbreviations, echo=FALSE}
abbreviations <- readr::read_delim(col_names = FALSE, delim = ";", trim_ws = TRUE, file = "
  AA; Aortic aneurysm
  AD;	Aortic dissection
  CDM;	Common Data Model
  OMOP;	Observational Medical Outcomes Partnership
  OHDSI;	Observational Health Data Science and Informatics
  RxNorm;	US-specific terminology that contains all medications available on the US market
  SNOMED;	Systematized Nomenclature of Medicine
  UTI;	Urinary Tract Infection
  ",
  show_col_types = FALSE)
tab <- knitr::kable(abbreviations, col.names = NULL, linesep = "", booktabs = TRUE)
if (knitr::is_latex_output()) {
  tab %>% kable_styling(latex_options = c("striped", "hold_position"), 
                        font_size = latex_table_font_size)
} else {
  tab %>% kable_styling(bootstrap_options = "striped")
}
```

\clearpage

# Responsible Parties

## Investigators

```{r parties, echo=FALSE}
parties <- readr::read_delim(col_names = TRUE, delim = ";", trim_ws = TRUE, file = "
  Investigator; Institution/Affiliation
  Jack Janetzki*;	University of South Australia
  Nicole Pratt*;	University of South Australia
  Seng Chan You*;	Yonsei University
  Seonji Kim;	Yonsei University
  Jung Ho Kim*;	Yonsei University
  Jung Ah Lee;	Yonsei University
  Patrick Ryan*;	Columbia University, Janssen
  ",
  show_col_types = FALSE)
tab <- kable(parties, booktabs = TRUE, linesep = "") %>%
  column_spec(1, width = "10em") %>%
  column_spec(2, width = "35em") %>%
  footnote(general = "* Principal Investigator", general_title = "")
if (knitr::is_latex_output()) {
  tab %>% kable_styling(latex_options = c("striped", "hold_position"),
                        font_size = latex_table_font_size)
} else {
  tab %>% kable_styling(bootstrap_options = "striped")
}
```

## Sponsor

This study is undertaken within Observational Health Data Sciences and Informatics (OHDSI), an open collaboration.

\clearpage

# Abstract

**Question** 
Are fluoroquinolones linked to increased risk of aortic aneurysm or dissection?

This distributed network analysis will describe the incidence and time-to event of aortic aneurysm and aortic dissection following exposure to fluoroquinolone antibiotics. The study will also  compare the risk of AA/AD  between individual fluoroquinolone antibiotics with other individual antibiotics used to treat the same infection (head-to-head comparisons). We will also compare the risk of AA/AD between fluoroquinolone antibiotics as a class with other individual antibiotics used to treat the same infection.  

**Background and Significance**: 

Fluoroquinolone antibiotics are a broad-spectrum class of antibiotics which are used to treat urinary tract infections, pneumonia, gastroenteritis, epididmyo-orchitis, prostatitis or bone and joint infections.  Fluoroquinolone use is rising internationally; approximately 7.81 billion doses of fluoroquinolones have been administered over the last decade. Whilst fluoroquinolone antibiotics are generally well tolerated, neurological and cardiovascular adverse events have been identified in post-market surveillance studies. Additionally in rare cases (0.1-1%), use of fluoroquinolones has been associated with risk of aortic aneurysm or dissection. The pharmacological mechanism between fluoroquinolone use and aortic aneurysm or dissection occurrence is not fully understood. However, the pathogenesis of this adverse event appears to involve extracellular matrix degradation in the aorta mediated by matrix metalloproteinases, dysfunction and apoptosis of vascular smooth muscle cells in the aorta and also increased concentrations of pro-inflammatory cytokines at the site contributing to remodelling of the aorta.  

Subsequent to post-market surveillance, international medicine regulators have issued safety warnings regarding the risk of aortic aneurysm or dissection (AA/AD) with fluoroquinolone antibiotics (FQ) antibiotics. Whilst there have been numerous warnings from regulators about the association of fluoroquinolone antibiotics with risk of aortic aneurysms or dissections, the quality of the evidence underpinning this association is moderate. The most recent meta-analysis of four observational studies found that people who used fluoroquinolone antibiotics were at an increased risk of aortic diseases compared to those who used other antibiotics (adjusted odds ratio 2.10; 95% CI 1.65-2.68)(1) and other studies have shown that risk of aortic dissection appears highest within 30 days of taking a fluoroquinolone and the risk remains elevated for a year(2). Prior studies supporting the risk are limited by inconsistencies in study design, approaches to identifying exposures and outcomes, and methods used to address confounding. 

***Study Aims/Objectives**: 

This study is a multinational cohort study which will estimate the risk of AA/AD following FQ exposure compared to other comparable antibiotics for urinary tract infection (UTI).

More specifically we will:

A. Characterizs the incidence and time-to event of aortic events following fluoroquinolone exposure 

B. Estimate the comparative safety of fluoroquinolones versus other antibiotics indicated for urinary tract infections


***Study Description**:

A large-scale  distributed network study will be undertaken via the OHDSI network. The selection of data sources will be based on their comprehensive population coverage, extensive historical data, diverse data capture processes, voluntary participation, and the availability of verified outcome information, ensuring robust and reliable research. 

***Population**: 

Participants will be patients 35 years or older initiating antibiotic treatment for UTI 

People taking systemic fluoroquinolones for treatment of urinary tract infections (UTI) 



* **Comparators**:
1. Trimethoprim with or without sulfamethoxazole (TMP)
2. Cephalosporins (CPH)



* **Outcomes**:
The primary outcome will be occurrence of AA/AD within 60 days after exposure. Secondary outcomes will be first occurrence of the individual outcome AA or AD.

Cox proportional hazards models will be used to estimate the risk of each outcome after 1:1 propensity score (PS) matching. All results will be calibrated based on the results from negative control outcomes. A series of pre-defined diagnostics will be employed to evaluate the analytical method performance. Only results of analyses passing all diagnostics will be reported. Hazard ratios (HRs) will be pooled across databases using Bayesian random effects meta-analysis. 


* **Design**: 

The study will be a retrospective cohort study. 

* **Timeframe**: 
January 2010 and December 2019. 

\clearpage

# Amendments and Updates

```{r amendments, echo=FALSE}
amendments <- readr::read_delim(col_names = TRUE, delim = ";", trim_ws = TRUE, file = "
  Number; Date; Section of study protocol; Amendment or update; Reason
",
  show_col_types = FALSE)
tab <- kable(amendments, booktabs = TRUE, linesep = "")
if (knitr::is_latex_output()) {
  tab %>% kable_styling(latex_options = c("striped", "hold_position"), font_size = latex_table_font_size)
} else {
  tab %>% kable_styling(bootstrap_options = "striped")
}
```

# Milestones

```{r dates, echo=FALSE}
dates <- readr::read_delim(col_names = TRUE, delim = ";", trim_ws = TRUE, file = "

The SOS challenge tutorial schedule is described below: 
•	Mar 28th 2023: Initiate network study, identifying data partners
•	Apr 4th 2023: Data diagnostics (data partners to share data diagnostics results)
•	Apr 11th 2023: Phenotype development
•	Apr 18th 2023: Phenotype evaluation (complete candidate phenotypes and data partners to share phenotype diagnostic results)
•	Apr 25th 2023: Create analysis specifications (finalize analytic package)
•	May 2nd 2023: Network execution (Data partners to run analytic package)
•	May 9th 2023: Study diagnostics (data partners to share study diagnostic results)
•	May 14 2023: Evidence synthesis
•	May 23 2023: Interpretation of results by investigators
•	May 2023 to June 204: Manuscript drafting and writing 
")

tab <- kable(dates, booktabs = TRUE, linesep = "")

if (knitr::is_latex_output()) {
  tab %>% kable_styling(latex_options = c("striped", "hold_position"), font_size = latex_table_font_size)
} else {
  tab %>% kable_styling(bootstrap_options = "striped")
}
```

# Rationale and Background

Fluoroquinolone antibiotics (FQ) are broad-spectrum antibiotics that treat a variety of infections, including urinary tract infections (UTIs). Although FQs are generally well tolerated, a recent meta-analysis based on five observational post-market studies concluded that FQ use doubled the risk of incident aortic diseases. While aortic aneurysm (AA) and aortic dissections (AD) are rare, they are fatal in 65-90% of cases, especially when an AA ruptures. 

In response to safety concerns raised by observational studies and adverse events reported to regulatory bodies internationally, warnings were issued regarding the risk of AA/AD with FQ antibiotics. Studies ahve shown that the use of outpatient FQs has decreased following these safety warnings in some but not all settings, with the United Kingdom Medicines and Healthcare products Regulatory Agency recently issuing further restrictions on prescribing. Patients have also shared concerns about the risk profile of FQs in public hearings. 

Recent studies have suggested that the previous studies reporting associations between FQs and some adverse drug reactions (ADRs) may have been affected by indication bias or surveillance bias. A meta-analysis of the four studies listed in the U.S. Federal Drug Administration (FDA) Drug Safety Communication regarding the association of FQs with AA/AD noted that confounding by indication may have influenced results and only one study included an active comparator. 


# Study Objectives
Given the conflicting results of previous studies, the recognised efficacy of FQs and the serious nature of AA/AD complications, we will conduct a large-scale distributed network study to estimate the risk of AA/AD after FQ exposure for treatment of UTI compared to other antibiotics. We will employ best-practice methodology to minimize potential systematic bias and use objective diagnostics to evaluate the analytical method performance. 

# Research Methods


## Study Design

A new-user cohort study will be conducted to estimate the comparative risk of AA/AD with new use of FQs compared to comparator antibiotics. 


## Data Sources

We will conduct a distributed network analysis across the Observational Health Data Science and Informatics (OHDSI) Data Network. This study will utilise routinely-collected health care data which has been mapped to the Observational Medical Outcomes Partnership (OMOP) Common Data Model (CDM).

The study will be conducted using data from real world data sources that have been mapped to the OMOP Common Data Model in collaboration with the Observational Health Data Sciences and Informatics (OHDSI) and European Health Data and Evidence Network (EHDEN) initiatives. The OMOP Common Data Model (https://ohdsi.github.io/CommonDataModel/) includes a standard representation of health care trajectories (such as information related to drug utilization and condition occurrence), as well as common vocabularies for coding clinical concepts, and enables consistent application of analyses across multiple disparate data sources (Voss et al., 2015). 

We intend to study data with different data source provenance (e.g., electronic medical records, insurance claims) as well data representing different populations (privately insured employees or patients with limited income) and geographies. This study will be run on datasets that have been converted to the Observational Medical Outcomes Partnership (OMOP) Common Data Model (CDM) 

We intend to study data with different data source provenance (e.g., electronic medical records, insurance claims) as well data representing different populations (privately insured employees or patients with limited income) and geographies. This study will be run on datasets that have been converted to the Observational Medical Outcomes Partnership (OMOP) Common Data Model (CDM) 

All participating sites and data partners will obtain approval or exemption from their institutional review boards prior to participating in the study. 

##Study time period
The study will utilize data from 1 January 2010 to 31 December 2019. The study period will be restricted to end of 2019 to avoid the potential impact of COVID-19 pandemic on patterns in antibiotic use. 

## Study Population

Study population: all subjects in the database will be included who meet the following criteria (note the index date is the start of the first exposure to fluoroquinolone or active comparator): 

Patients will be included if at index date they are aged 35 years or older, have at least 365 days of prior observation in the database, have a recorded condition occurrence of UTI on or within 7 days, and were not hospitalized for any reason within 7 days prior to the index date. Patients could have no diagnosis of aortic aneurysm or dissection preceding the index date.

The study population will be defined as:  

All people included in this analysis must be diagnosed with urinary tract infection. For all antibiotics, we restrict to people with a diagnosis of Urinary Tract Infection, a major indication for the treatment of interest. To minimize confounding by indication we will restrict analysis to patients treated in an outpatient setting for UTIs, which are commonly localised, well-characterized infections with a simliar severity profile across presentations. 


Three exposure cohorts will be generated: 
(1) FQ exposure, 
(2) TMP exposure, or
(3) CPH exposure. 

Only systemic exposure from medications administered orally or via injection will be included in the exposure definitions. TMP and CPH were selected as comparator antibiotics as they are generally recommended in treatment guidelines for UTI. 



Included fluoroquinolones and concept code: 

Concept ID    Concept name 
1721543       norfloxacin 
35198003      pazufloxacin mesilate 
1747032       grepafloxacin 
19041153      temafloxacin 
1592954       delafloxacin 
1716721       gemifloxacin 
923081        ofloxacin 
19050750      fleroxacin 
1712549       trovafloxacin 
35197938      garenoxacin mesilate hydrate 
36878831      nadifloxacin 
40161662      besifloxacin 
43009030      tosufloxacin tosylate hydrate 
1743222       enoxacin 
35834909      lascufloxacin hydrochloride 
35198165      sitafloxacin hydrate 
1789276       gatifloxacin 
1797513       ciprofloxacin 
1742253       levofloxacin 
1716903       moxifloxacin 
35197897      prulifloxacin 
19027679      pefloxacin 
1733765       sparfloxacin 
1707800       lomefloxacin 

## Exposure Comparators

1. Trimethoprim with or without sulfamethoxazole (TMP)


Concept ID     Concept Name 
40081374       sulfamethoxazole / trimethoprim Injectable Solution 
36882762       Sulfamethoxazole / Trimethoprim Injectable Suspension 
40220482       sulfamethoxazole / trimethoprim Injection 
35153537       Sulfamethoxazole / Trimethoprim Oral Granules 
40147374       sulfamethoxazole / trimethoprim Oral Solution 
40081379       sulfamethoxazole / trimethoprim Oral Suspension 
40081388       sulfamethoxazole / trimethoprim Oral Tablet 
1705674        trimethoprim 

2. Cephalosporin (CPH)


Concept ID      Concept Name 
40798709        Cefacetrile 
1796435         cefixime 
40798704        Cefmenoxime 
19072255        cefmetazole 
43008993        cefminox sodium 
19028286        cefodizime 
19072857        cefonicid 
1773402         cefoperazone 
19028288        ceforanide 
1774470         cefotaxime 
1774932         cefotetan 
19051271        cefotiam 
1775741         cefoxitin 
43009045        cefpiramide sodium 
1749008         cefpodoxime 
1738366         cefprozil 
43009083        cefroxadine 
19051345        cefsulodin 
1776684         ceftazidime 
35198137        cefteram pivoxil 
43008994        ceftezole sodium 
1749083         ceftibuten 
1777254         ceftizoxime 
1777806         ceftriaxone 
1778162         cefuroxime 
1786621         cephalexin 
19052683        cephaloridine 
19086759        cephalothin 
19086790        cephapirin 
1786842         cephradine 
43009087        flomoxef sodium 
1708100         loracarbef 
19126622        moxalactam 
1768849         cefaclor 
1769535         cefadroxil 
19070174        cefamandole 
19070680        cefatrizine 
40798700        Cefazedone 
1771162         cefazolin 
43009082        cefbuperazone sodium 
43009044        cefcapene pivoxil hydrochloride hydrate 
1796458         cefdinir 
1747005         cefditoren 
19028241        cefetamet 

## Outcomes {#outcomes}

The primary outcome will be diagnosis of either AA or AD during a hospital or emergency department visit in the 60 days after index date. The secondary endpoints included the occurrence of each outcome separately. Patients will be excluded if they have a recorded outcome event in the 365 days prior to index date. A time at risk window of 60 days after treatment initiation will be employed as the primary analysis as this is consistent with previous studies. We will also vary the time at risk window as: 30, 60, 90 and 365 days as sensitivity analyses. 

Details of the outcome definitions are provided below: 

###Aortic Aneurysm
ICD-9-CM:
441;441.0; 441.1; 441.2; 441.3; 441.3; 441.4; 441.5; 441.6; 441.7; 441.9

ICD-10:
I71.0; I71.1; I71.2; I71.3; I71.4; I71.5; I71.6; I71.8; I71.9

CPT-4:
001T; 0002T' 0033T; 0034T; 0035T; 0036T; 0039T; 0078T; 0079T; 0080T; 0081T; 33720; 33877; 33880; 33881; 34800; 34802; 34803; 34804; 34805; 34813; 34830; 34831; 34832; 35081; 35082; 35091; 35092; 35102; 35103; 75952; 75953; 9001F; 9003F; 9004f

ICD-9 PROCEDURE:
39.71; 39.73

Positive Predictive Value was calculated for this outcome based on hospitalization or emergency room (ER) visits at any diagnosis position based on the ICD-10 code in the single Korean tertiary hospital, Severance Hospital (Yonsei University Heatlh System). PPV, % (n) = 97 (97/100)

###Aortic dissection
ICD-9-CM: 
441.0; 441.00; 441.01; 441.02; 441.03

ICD-10:
I71.0

PPV= 79% (N= 79/100)

###Negative control outcomes
Negative control outcomes are known to have no association with the target (fluoroquinolone) or comparator (trimethoprim or cephalosporin) cohorts, such that we can assume the true relative risk between the two cohorts is 1. Negative control outcomes for this study will be selected based on a similar process to that outlined by Voss et al. Once potential negative control candidates are selected, a manual clinical review will be performed to exclude any pairs that may have a causal relationshipu or are similar to the study outcome. 

The top 50 outcomes based on their prevalence will be selected. 

After doing, so the final list of 50 negative outcomes to be used in the study are: 


OMOP Concept ID     Outcome Name 

4170145             Absence of lung 

4093531             Absence of toe 

4092879             Absent kidney 

42539582            Acquired absence of genital organ 

434170              Atypical squamous cells of undetermined significance on cervical Papanicolaou smear 

4067069             Callosity 

4213540             Cervical somatic dysfunction 

443570              Cervicovaginal cytology: Low grade squamous intraepithelial lesion 

201613              Chronic nonalcoholic liver disease 

43021250            Complication associated with orthopedic device 

46269889            Complication due to Crohn's disease 

42537730            Coronary artery graft present 

436233              Delayed milestone 

438021              Disorder due to and following fracture of upper limb 

192367              Dysplasia of cervix 

4062791             Endocrine, nutritional and metabolic disease complicating pregnancy, childbirth and puerperium 

200775              Endometrial hyperplasia 

374358              Excess skin of eyelid 

4059015             Falls 

4264617             Foot-drop 

4201388             Gastrostomy present 

4166231             Genetic predisposition 

4295287             Hypercoagulability state 

196473              Hypertrophy of uterus 

443447              Iatrogenic hypotension 

4344500             Impingement syndrome of shoulder region 

441417              Incoordination 

4168222             Intra-abdominal and pelvic swelling, mass and lump 

196168              Irregular periods 

439795              Minimal cognitive impairment 

40480893            Nonspecific tuberculin test reaction 

438130              Opioid abuse 

4022076             Patient dependence on care provider 

4141640             Perimenopausal disorder 

437092              Physiological development failure 

4012231             Poor stream of urine 

46286594            Problem related to lifestyle 

443274              Psychostimulant dependence 

436246              Reduced libido 

43021237            Secondary erectile dysfunction 

4052226             Sequelae of injuries of lower limb 

4125590             Slurred speech 

36713918            Somatic dysfunction of lumbar region 

4002818             Spasm of back muscles 

4008710             Stenosis due to any device, implant AND/OR graft 

4201387             Tracheostomy present 

42538119            Transplanted heart valve present 

444074              Victim of vehicular AND/OR traffic accident 

195603              Vulval and/or perineal noninflammatory disorders 

4216670             Worried well 



## Analysis

Characterizing the cohort:  

All analyses will be performed using code developed for the OHDSI Methods library. The code for this study can be found at A diagnostic package, built off the OHDSI Cohort Diagnostics library, is included in the base package as a preliminary step to assess the fitness of use of phenotypes on your database. If a database passes cohort diagnostics, the full study package will be executed. Baseline covariates will be extracted using an optimized SQL extraction script based on principles of the Feature Extraction package () to quantify Demographics (Gender, Prior Observation Time, Age Group), Condition Group Eras and Drug Group Eras (at the above-listed time windows). Additional cohort-specific covariates will be constructed using OMOP standard vocabulary concepts. 



Number and proportion of persons with feature variables during time-at-risk windows will be reported by target cohort and specific stratifications. Standardized mean differences (SMD) will be calculated when comparing characteristics of study cohorts, with plots comparing the mean values of characteristics for each of the features (with the color indicating the absolute value of the standardized difference of the mean). 



Population-level estimation:  

Proportional hazard models will be used to assess the hazard ratios between the two exposure cohorts for each condition of interest and for each setting.  

Adjustment for baseline confounders will be done using propensity scores. First a propensity model will be fitted and used to create propensity scores (PS). These PS will be used ot match the treatment and comparator cohorts, and the proportional hazards outcome models will be conditioned on the matched sets of strata respectively.  

The analysis will be performed as an intention to treat where exposure is considered from index date until end of the observation (until the occurrence of outcome, death or end of study).  

Negative controls: as described above. The hazard ratios computed for these negative controls will be used to evaluate residual bias and compute calibrated p-values for the outcomes of interest. 

Cox proportional hazards will be used to estimate the hazard ratio of each outcome for each time at risk window in each data source after propensity score (PS) matching. Propensity score models derivation is described below in 'Covariate balance'. 

A Bayesian random-effects meta-analysis will be conducted to combine each sites hazard ratio estimate into a single aggregated hazard ratio using non-normal likelihood approximations to avoid bias due to small or zero counts. 

##Study diagnostics
Analyses using observational data can produce misleading estimates as a result of study design and analytic choices. We will implement several study diagnostics using pre-specified decision thresholds to evaluate the reliability of our analyses. These study diagnostics will assist with understanding the risk of bias and generalizability of the estimates generated in each database across the network of databases involved in the study. Only results that passed the pre-specified thresholds set for each diagnostics will contribute to the meta-analytic estimates. 

###Covariate balance
We will implement propensity score matching to account for confounding between treatment groups. Propensity score models will include covariates such as age, race, medications, medical conditions, procedure exposure, medical exposure, and laboratory values. The number of covariates to be included will be between 10,000 to 100,000. Covariate use behaviours over the 30 to 365 days prior to index date will be utliised for PS model calculation.

Large-scale regularized regression will be used to fit the propensity model. Separate PS models will be constructed for each comparator. Patients will be matched 1:1 between the target and comparator cohorts. 

To determine whether the PS matching is sufficient to balance baseline patient characteristics we will calculate the standardised mean difference between treatment cohorts after propensity score matching. Covariate balance diagnostic will be achieved if all SMDs of predefined covariates below are less than 0.1.  


Age group: Age was grouped in 5-year intervals 

Gender 

Race 

Ethnicity 

Medical history: Acute respiratory disease, Attention deficit hyperactivity disorder, Chronic liver disease, Chronic obstructive lung disease, Crohn's disease, Dementia, Depressive disorder, Diabetes mellitus, Gastroesophageal reflux disease, Gastrointestinal hemorrhage, Human immunodeficiency virus infection, Hyperlipidemia,Hypertensive disorder, Lesion of liver, Obesity, Osteoarthritis, Pneumonia, Psoriasis, Renal impairment, Rheumatoid arthritis, Schizophrenia, Substance abuse, Ulcerative colitis, Viral hepatitis C, Visual system disorder, Atrial fibrillation, Cerebrovascular disease, Coronary arteriosclerosis, Heart disease, Heart failure, Ischemic heart disease, Peripheral vascular disease, Pulmonary embolism, Venous thrombosis, Hematologic neoplasm, Malignant lymphoma, Malignant neoplasm of anorectum, Malignant neoplastic disease, Malignant tumor of breast, Malignant tumor of colon, Malignant tumor of lung, Malignant tumor of urinary bladder, Primary malignant neoplasm of prostate 

Medication use: Agents acting on the renin-angiotensin system, Antibacterials for systemic use, Antidepressants, Antiepileptics, Antiinflammatory and antirheumatic products, Antineoplastic agents, Antipsoriatics, Antithrombotic agents, Beta blocking agents, Calcium channel blockers, Diuretics, Drugs for acid related disorders, Drugs for obstructive airway diseases, Drugs used in diabetes, Immunosuppressants, Lipid modifying agents, Opioids, Psycholeptics, Psychostimulants, Agents used for ADHD and nootropics 

Comorbidity index: Charlson Comorbidity Index, CHA2DS2-Vasc score, Diabetes Complications Severity Index 

###Clinical equipoise
Empirical equipoise will be assessed by determining the overlap in preference score distribution between the target and comparator cohorts. The preference score distribution is a transformation of the propensity score. High overlap, and greater equipoise ensures that results will be generalisable back to the original cohort. Good equipoise means that even a large propensity score model could not discriminate between preference for two treatments. This is similar to randomised clinical trials where study participants have the same probability of receiving either intervention in a 1:1 randomised trial regardless of their characteristics. at the clinical equipoise diagnostic was achieved if at least 20% of matched patients had a preference score between 0.3 and 0.7.  

###Systematic error
To determine systematic error or residual bias (due to study design and analytic choices), we will estimate systematic error by employing negative controls. Negative controls (as described above) are determined as outcomes which are assumed to not be associated with either the target or comparator treatment. We will include 50 negative controls in our analysis. For each of these outcomes we will implement the analysis as per the primary analysis and we compare the estimated result of the analysis against the “known truth”, in this case we expect the true realative risk to be 1. Positive controls will not be utilised as there are no medicines which are known to increase risk of aortic aneurysm or dissection. 

Overall systematic error will be calculated as the Expected Absolute Systematic Error (EASE) score. EASE will be calculated by first fitting the systematic error distribution across the set of negative control outcomes and then taking the aboslute expected value of the distribution.

If EASE score is low, then calibrated and uncalibrated estimates (hazard ratios, p-values, confidence intervals) will be similar indicating little to no systematic error or unmeasured confounding that is not being accounted for in the results of the study. The systematic bias diagnostic will be achieved if the EASE value was less than 0.25. 

Study diagnostics: Overall an analysis was considered to have passed study diagnostics if SMDs of predefined covariates were less than 0.1, equipoise was greater than 0.2 and EASE was less than 0.25. 


# Sample Size and Study Power {#sample-size}

Sample size and study power will be calculated based on the number of patients included from the databases passing diagnostics and patients prescribed exposure drug (FQ) and comparators (CPH or TMP) for treatment of UTI. 

# Strengths and Limitations {#strengths-limitations}

##Strengths
This large-scale distributed network study wil' use a rigorous design with objective diagnostics to reduce bias and confounding. The study design ensures transparency, reproducibility and reliability of results. 

To minimize confounding by indication, we will restrict the analysis to patients treated in an outpatient setting for UTIs, which are commonly localized, well-characterized infections with a similar severity profile across presentations.

We will use this detailed protocol with standardized definitions of exposures, outcome measures, study design and approach to address confounding. 

We will also use a standardized analytic package to generate the analytic results within each of the participating databases. 

Our approach also mirrors that of a hypothetical randomized controlled trial by employing a new-user, active comparator design with clearly defined indications and we will implmement an outcome-agnostic, data-driven PS model, modelled using large-scale regularised regression. 

We will also apply the suite of predefined objective diagnostics as described above to esnure the quality of evidence focusing on sufficient covaraiate balance, overlap in preference score distributions, and minimal systematic error as evluated by a comprehensive set of negative control outcomes. 

Only analyses that pass all objective diagnostic tests will be included in the meta-analysis. 

##Limitations
The definition of AA/AD will rely on ICD-9-CM and ICD-10-CM codes rather than imaging modalities. We have however validated our specific diagnosis codes in one Korean tertiary hospital with high PPV for the outcomes of investigation. 

Results may not be generalizable to other indications as we only examined FQs used to treat UTIs in the outpatient setting.

Antibiotics are usually used for a short period of time to treat UTIs, therefore, we are unable to investigate the effects of long-term cumulative use of FQs in this study. 

Regulatory warnings about the use of FQs during the study may affect results, however the majority of the study period is before such warnings were made. 

We are unable to consider the dose of FQ or comparator antibiotics in this study. As the critiera for antibiotic dosage and duration are well established in outpatient UTIs, it is likely that variation in dose would be small, however we are unable to determine whether there is a dose-response relationship. 

# Protection of Human Subjects

The study uses only de-identified data. Confidentiality of patient records will be maintained at all times. Data custodians will remain in full control of executing the analysis and packaging results. There will be no transmission of patient-level data at any time during these analyses. Only aggregate statistics will be captured. Study packages will contain minimum cell count parameters to obscure any cells which fall below allowable reportable limits. All study reports will only contain aggregated data and will not identify individual patients or physicians.  

# Management and Reporting of Adverse Events and Adverse Reactions

N/A

# Plans for Disseminating and Communicating Study Results

Share results and study design at OHDSI conferences.

Share primary results in high-impact journal. 

\clearpage

# References {-}

Dai XC, Yang XX, Ma L, Tang GM, Pan YY, Hu HL. Relationship between fluoroquinolones and the risk of aortic diseases: a meta-analysis of observational studies. BMC Cardiovasc Disord. 2020;20(1):49. 

Yu X, Jiang DS, Wang J, Wang R, Chen T, Wang K, Cao S, Wei X. Fluoroquinolone Use and the Risk of Collagen-Associated Adverse Events: A Systematic Review and Meta-Analysis. Drug Saf. 2019;42(9):1025-1033. 

Voss EA, Boyce RD, Ryan PB, van der Lei J, Rijnbeek PR, Schuemie MJ. Accuracy of an automated knowledge base for identifying drug adverse reactions. J Biomed Inform 2017; 66: 72–81.

<div id="refs"></div>

\clearpage

\centerline{\Huge Appendix}

# (APPENDIX) Appendix {-}

# Exposure Cohort Definitions

```{r appendix,  echo=FALSE, results="asis", warning=FALSE, message=FALSE}
# source("https://raw.githubusercontent.com/ohdsi-studies/LegendT2dm/master/R/PrettyOutput.R")
# printCohortDefinitionFromNameAndJson(name = "Template Cohort (New-users of DDP4 inhibitors)",
#                                      json = SqlRender::readSql("templateCohort.json"),
#                                      withConcepts = TRUE)
```

# Outcome Cohort Definitions

# Negative Control Concepts {#negative-controls}